### 2024-03-10 paper/paper.qmd
update

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Hailey Jang
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-03-13 scripts/00-simulate_data.R
Update scripts with data

@@ -1,19 +1,38 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Generate simulated data for annual deaths, categorized by age and sex
+# Author: Hailey Jang
+# Date: 12 March 2024
+# Contact: hailey.jang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: NA
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
 
 #### Simulate data ####
-# [...ADD CODE HERE...]
+# Set the seed for reproducibility
+set.seed(93)
+
+# Simulate data
+years <- rep(seq(2017, 2023, 1), each = 5)
+age <- sample(c("40-59", "20-39", "60+", "<20"), size = 35, replace = TRUE, prob = c(0.2, 0.2, 0.1, 0.5))
+sex <- sample(c("Female", "Male"), size = 35, replace = TRUE, prob = c(0.35, 0.65))
+death_counts <- round(rpois(length(years), lambda = 9.5))
+
+simulated_data <- data.frame(
+  Year_of_death = years,
+  Age_group = age,
+  Gender = sex,
+  Count = death_counts
+)
+
+### Save simulated data ###
+write_csv(simulated_data, "data/raw_data/Homeless_deaths_by_demographics.csv")
+
+
+
+
 
 
 

### 2024-03-13 scripts/01-download_data.R

@@ -1,26 +1,29 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from opendatatoronto
+# Author: Hailey Jang
+# Date: 12 March 2024
+# Contact: hailey.jang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
 
 #### Workspace setup ####
 library(opendatatoronto)
 library(tidyverse)
-# [...UPDATE THIS...]
 
 #### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
+# Got the data from OpenDataToronto
+# Code based on https://open.toronto.ca/dataset/deaths-of-people-experiencing-homelessness/
 
-
+# Read CSV file 
+homeless_death <-
+  read_csv(
+    file = "https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/a7ae08f3-c512-4a88-bb3c-ab40eca50c5e/resource/ceafdcdf-5f41-4ba5-9b69-f4c8e581ab6e/download/Homeless%20deaths%20by%20cause.csv", 
+    show_col_types = FALSE
+  )
 
 #### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
 
-         
+write.csv(
+  x = homeless_death, 
+  file = "data/raw_data/Homeless_deaths_by_cause.csv"
+)
+

### 2024-03-13 scripts/02-data_cleaning.R

@@ -1,44 +1,27 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: To filter the dataset, focusing on non-suicide related mortality records
+# Author: Hailey jang 
+# Date: 11 March, 2024
+# Contact: hailey.jang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 #### Workspace setup ####
 library(tidyverse)
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+dataset <- read_csv("data/raw_data/Homeless_deaths_by_cause.csv")
 
+# Rename year of death column
+names(dataset)[names(dataset) == 'Year of death'] <- 'Year_of_death'
+
+# Clean data to remove unknown values and isolate causes of deaths relating to suicide
 cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
+  dataset |>
+  filter(Cause_of_death != "Suicide", Age_group != "Unknown", Gender != "Unknown") |>
[…]

### 2024-03-13 scripts/03-test_data.R

@@ -1,15 +1,39 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Validate simulation outputs and data cleaning 
+# Author: Hailey jang
+# Date: 11 March, 2024
+# Contact: hailey.jang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+analyzed_data <- read_csv("data/analysis_data/cleaned_data.csv")
 
 #### Test data ####
+# Verify the integrity of four distinct variables
+analyzed_data %>% 
+  unique() %>% 
+  length() == 4
+
+# Validate data types of each variable
+is.numeric(analyzed_data$Year_of_death)
+is.character(analyzed_data$Age_group)
+is.character(analyzed_data$Gender)
+is.numeric(analyzed_data$Count)
+
+# Confirm that the highest recorded deaths in any category are within expected bounds
+max(analyzed_data$Count) <= 75
+
+# Ensure that the lowest count of deaths is not less than zero
+min(analyzed_data$Count) >= 0 
+
+# Validate the presence of all specified age groups
+unique(analyzed_data$Age_group) == c("40-59", "20-39", "60+", "<20")
+
+# Check for the inclusion of all expected years in the data
+unique(analyzed_data$Year_of_death) == c('2017', '2018', '2019', '2020', '2021', '2022', '2023')
+
+# Ensure gender data contains only the specified categories
+unique(analyzed_data$Gender) == c('Female', 'Male')
+

### 2024-03-13 scripts/04-model.R

@@ -1,11 +1,9 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Conduct statistical modeling
+# Author: Hailey jang 
+# Date: 11 March, 2024
+# Contact: hailey.jang@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 
 #### Workspace setup ####
@@ -13,25 +11,29 @@ library(tidyverse)
 library(rstanarm)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+modeling_data <- read_csv("data/analysis_data/cleaned_data.csv")
 
 ### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+first_model <- stan_glm(
+  formula = Count ~ Gender + Age_group,
+  data = modeling_data,
+  family = gaussian(),
+  prior = normal(0, 23, autoscale = TRUE),
+  prior_intercept = normal(0, 1, autoscale = TRUE),
+  prior_aux = exponential(0.33, autoscale = TRUE),
+  seed = 80
+)
 
 
 #### Save model ####
+# Save the constructed model for future reference
+saveRDS(first_model, file = "models/bayesian_model.rds")
+
 saveRDS(
   first_model,
   file = "models/first_model.rds"
 )
 
+# Notification of script completion
+message("first_model has been successfully created and saved.")
 

### 2024-03-13 scripts/04-model.R
update

@@ -27,8 +27,6 @@ first_model <- stan_glm(
 
 #### Save model ####
 # Save the constructed model for future reference
-saveRDS(first_model, file = "models/bayesian_model.rds")
-
 saveRDS(
   first_model,
   file = "models/first_model.rds"

### 2024-03-14 paper/paper.qmd
Update data

@@
   - Hailey Jang
-thanks: "Code and data are available at: LINK."
+thanks: "Code and data are available at: https://github.com/Hailey-Jang/Suicide-and-Homelessness.git"
 date: today

### 2024-03-14 scripts/02-data_cleaning.R

@@ -11,17 +11,103 @@ library(tidyverse)
 #### Clean data ####
 dataset <- read_csv("data/raw_data/Homeless_deaths_by_cause.csv")
 
-# Rename year of death column
+# Rename the year of death column
 names(dataset)[names(dataset) == 'Year of death'] <- 'Year_of_death'
 
 # Clean data to remove unknown values and isolate causes of deaths relating to suicide
 cleaned_data <-
   dataset |>
   filter(Cause_of_death != "Suicide", Age_group != "Unknown", Gender != "Unknown") |>
-  # Pipe this to include only year of death, age group, gender, and count.
+  # Pipe this to include only the year of death, age group, gender, and count.
   select(Year_of_death, Age_group, Gender, Count)
 
 head(cleaned_data, n=10)
 
 #### Save data ####
 write_csv(cleaned_data, "data/analysis_data/cleaned_data.csv")
+
+### Data for sex###
+
+# Initialize a list to store the sum of deaths of Male
+Male_deaths_summary <- list()
+
+for (year in c(2017,2018,2019,2020,2021, 2022, 2023)) {
+  # Filter data for Male
+  year_data <- cleaned_data %>%
+    filter(Year_of_death == year, Gender == 'Male') %>%
+    select(Year_of_death, Gender, Count)
+  
+  print(head(year_data, n = 10))
+  
+  # Calculate the sum of deaths
+  Male_deaths_summary[[as.character(year)]] <- sum(year_data$Count)
+}
+
+# Initialize a list to store the sum of deaths of Female
+Female_deaths_summary <- list()
+
+for (year in c(2017,2018,2019,2020,2021, 2022, 2023)) {
+  # Filter data for Female
+  year_data <- cleaned_data %>%
+    filter(Year_of_death == year, Gender == 'Male') %>%
+    select(Year_of_death, Gender, Count)
+  
+  print(head(year_data, n = 10))
+  
+  # Calculate the sum of deaths
+  Female_deaths_summary[[as.character(year)]] <- sum(year_data$Count)
+}
+
+# Load the tibble package for creating tibbles
+library(tibble)
+
+# Create a tibble to organize the death data by year and gender
+gender_deaths_data <- tibble(
+  Year = rep(c('2017', '2018', '2019', '2020', '2021', '2022', '2023'), each = 2),
+  Gender = rep(c('Male', 'Female'), times = 7),
[…]

### 2024-03-14 paper/paper.qmd
update

--- 2024-03-10 paper/paper.qmd
@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "An Analytical Exploration of Homelessness and Suicide in Toronto"
 author: 
   - Hailey Jang
-thanks: "Code and data are available at: LINK."
+thanks: "Code and data are available at: https://github.com/Hailey-Jang/Suicide-and-Homelessness.git"
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper presents a study focused on addressing the critical issue of suicide among Toronto's homeless population, a significant public health challenge.
+We developed a risk stratification model using sociodemographic data to identify homeless individuals at high risk for suicide, employing a generalized linear model that incorporates factors like sex and age.
+The findings, derived from data collected by Toronto Public Health since January 2017, offer insights into the patterns and predictors of suicide attempts within this vulnerable group.
+The importance of this research lies in its potential to guide interventions, enabling more effective prevention strategies for suicide among the homeless, particularly in settings where immediate access to mental health professionals is limited."
 format: pdf
@@
 library(palmerpenguins)
+library(knitr)
+library(kableExtra)
 ```
@@
 
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
+Homelessness represents a significant public health crisis, with the intersection of inadequate housing and mental health issues exacerbating the risk of suicide among this population.
+Individuals experiencing homelessness face a 2- to 6-fold increased risk of suicide compared to the general population (Mark 2017).
+This alarming statistic not only highlights the vulnerability of homeless individuals but also underscores the urgent need for targeted interventions and preventive measures.
+This paper delves into the development of a risk stratification model designed to identify homeless individuals at an elevated risk of suicide.
+By integrating sociodemographic data within a generalized linear model, focusing on sex and age as primary factors, our research aims to provide a predictive tool for healthcare providers and social services.
+The motivation behind this study stems from the critical gap in existing research regarding effective, data-driven strategies for suicide prevention in the homeless community.
 
-The remainder of this paper is structured as follows. @sec-data....
+Our research methodology involved the analysis of comprehensive data collected by Toronto Public Health, starting from January 2017.
+This dataset not only provided a foundation for our model but also enriched our understanding of the prevalence and causes of suicide among the homeless.
+The outcomes of our study revealed significant patterns and predictors that can enhance the effectiveness of suicide prevention strategies, offering a new perspective on addressing this public health challenge.
+The significance of our findings extends beyond academic interest; they offer practical insights for policymakers, healthcare providers, and social workers who are directly involved in supporting the homeless population.
+By identifying high-risk individuals, our model can inform targeted interventions, potentially saving lives and allocating resources more efficiently.
+
+The structure of the paper is organized as follows: The data section delves into the broader context of the dataset, emphasizing the crucial aspects of measurement relevant to our study.
+The results section is comprehensive, encompassing various illustrative tables and graphs alongside detailed statistical analyses.
+Subsequently, the model section elucidates the setup and justification of our chosen modelling approach, ensuring clarity and transparency of our methodology.
+Lastly, the discussion articulates the study's contributions to our understanding of homelessness and suicide and outlines directions for future research, ensuring a thorough contemplation of the study's broader implications.
 
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+The study was performed using data from the City of Toronto’s **opendatatoronto** database portal, accessed through the ‘opendatatoronto’ package and processed using the statistical programming environment **R**.
+The **tidyverse** package facilitated the data and the **kableExtra** package was utilized for enhancing table presentations.
+Specific to this study, the LaTeX package **"float"** was implemented in the R markdown setting to ensure stable positioning of figures and tables, as detailed by the **"floatplacement{figure}{H}"** command.
+
+Initiated in January 2017, Toronto Public Health (TPH) embarked on a systematic record-keeping of homeless mortality to gain a clearer understanding of the prevalence and causative trends of these incidents.
+The dataset comprises variables such as the **year of death, cause of death, age group, gender**, and **number of deaths**.
+It classifies individuals into age categories spanning 20 years, starting from **"<20"** years to **"60+"** years, with gender recorded as either **"Male"** or **"Female"**.
+The dataset enumerates the deaths annually from **2017** through **2023**, categorizing the causes into distinct classifications like **"Accident", "Drug Toxicity", "Suicide"**, among others.
+
+Focusing on suicide-related fatalities, the dataset was refined to exclude entries marked as **"Unknown"** or **empty**.
+This filtration led to the construction of two specialized datasets: one delineating the yearly suicide death toll segregated by gender and another by age group.
+This reorganization necessitated aggregating individual counts from each report into a consolidated figure for these subgroups, facilitating a focused examination of suicide trends across different demographics over the years.
+
+Table @tbl-suicide-age-group presents the refined data, specifically spotlighting instances of suicide by age groups, and depicting the temporal and numerical specifics of each case.
 
@@
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+#| label: tbl-suicide-age-group
+#| tbl-cap: Summary of Suicides by Age Group
+#| echo: false
+
+# Loading the dataset
+suicide_by_age <- read_csv(here::here("data/analysis_data/age_group_deaths_data.csv"))
+
+suicide_by_age %>%
+  select(Year, Age_group, Deaths) %>%
+  head(8) %>%
+  kable(
+    col.names = c("Year", "Age Group", "Deaths"),
+    digits = 1,
+    format = "latex",
+    booktabs = TRUE,
+    linesep = ""
+  ) %>%
+
+  # Ensuring the table maintains its designated position
+  kable_styling(latex_options = "HOLD_position")
 ```
 
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+Table @tbl-suicide-gender-group presents the refined data, specifically spotlighting instances of suicide by gender, and depicting the temporal and numerical specifics of each case.
 
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
+#| label: tbl-suicide-gender-group
+#| fig-cap: Summary of suicides Rates by Gender
 #| echo: false
@@
 
-analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
+library(knitr)
 
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
+# Load the dataset containing gender-grouped suicide data
+gender_data <- read_csv(here::here("data/analysis_data/gender_deaths_data.csv"))
+
+# Refine the age-group dataset to focus on entries specific to suicides
+gender_data |> 
+  select(Year, Gender, Deaths) |>  # Select relevant columns
+  slice(1:8) |>  
+  kable(
+    col.names = c("Year", "Gender", "Deaths"),  
+    digits = 1,  
+    booktabs = TRUE,  # Use booktabs style for the table
+    linesep = "") |>  
+  kable_styling(latex_options = "HOLD_position")
 ```
-
-Talk way more about it.
 

### 2024-03-14 paper/paper.qmd
update

@@
 date: today
-date-format: long
 abstract: "This paper presents a study focused on addressing the critical issue of suicide among Toronto's homeless population, a significant public health challenge.
@@
 The importance of this research lies in its potential to guide interventions, enabling more effective prevention strategies for suicide among the homeless, particularly in settings where immediate access to mental health professionals is limited."
-format: pdf
+format:
+  pdf:
+    toc: true
     number-sections: true
@@
 library(knitr)
+library(here)
 library(kableExtra)
@@
 Homelessness represents a significant public health crisis, with the intersection of inadequate housing and mental health issues exacerbating the risk of suicide among this population.
-Individuals experiencing homelessness face a 2- to 6-fold increased risk of suicide compared to the general population (Mark 2017).
+Individuals experiencing homelessness face a 2- to 6-fold increased risk of suicide compared to the general population (@article1).
 This alarming statistic not only highlights the vulnerability of homeless individuals but also underscores the urgent need for targeted interventions and preventive measures.
@@
 
-The study was performed using data from the City of Toronto’s **opendatatoronto** database portal, accessed through the ‘opendatatoronto’ package and processed using the statistical programming environment **R**.
+The study was performed using data from the City of Toronto's **opendatatoronto** database portal, accessed through the 'opendatatoronto' package and processed using the statistical programming environment **R**.
 The **tidyverse** package facilitated the data and the **kableExtra** package was utilized for enhancing table presentations.
@@
 The dataset comprises variables such as the **year of death, cause of death, age group, gender**, and **number of deaths**.
-It classifies individuals into age categories spanning 20 years, starting from **"<20"** years to **"60+"** years, with gender recorded as either **"Male"** or **"Female"**.
+It classifies individuals into age categories spanning 20 years, starting from **"\<20"** years to **"60+"** years, with gender recorded as either **"Male"** or **"Female"**.
 The dataset enumerates the deaths annually from **2017** through **2023**, categorizing the causes into distinct classifications like **"Accident", "Drug Toxicity", "Suicide"**, among others.
@@
 
-Table @tbl-suicide-age-group presents the refined data, specifically spotlighting instances of suicide by age groups, and depicting the temporal and numerical specifics of each case.
+Table tbl-suicide-age-group presents the refined data, specifically spotlighting instances of suicide by age groups, and depicting the temporal and numerical specifics of each case.
 
@@
 
-Table @tbl-suicide-gender-group presents the refined data, specifically spotlighting instances of suicide by gender, and depicting the temporal and numerical specifics of each case.
+Table tbl-suicide-gender-group presents the refined data, specifically spotlighting instances of suicide by gender, and depicting the temporal and numerical specifics of each case.
 
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
 ## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
 
@@
 
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
-
-# Discussion
-
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
-
-\newpage
-
-\appendix
-
-# Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
-
-\newpage
-
 # References

### 2024-03-14 paper/paper.qmd
Update

@@
 library(kableExtra)
+library(readr)
+library(tidyr)
+library(dplyr)
 ```
@@
 
+## Data source
+
 The study was performed using data from the City of Toronto's **opendatatoronto** database portal, accessed through the 'opendatatoronto' package and processed using the statistical programming environment **R**.
 The **tidyverse** package facilitated the data and the **kableExtra** package was utilized for enhancing table presentations.
-Specific to this study, the LaTeX package **"float"** was implemented in the R markdown setting to ensure stable positioning of figures and tables, as detailed by the **"floatplacement{figure}{H}"** command.
+Specific to this study, the LaTeX package was implemented in the R markdown setting to ensure stable positioning of figures and tables.
+
+### Data measurment
 
 Initiated in January 2017, Toronto Public Health (TPH) embarked on a systematic record-keeping of homeless mortality to gain a clearer understanding of the prevalence and causative trends of these incidents.
-The dataset comprises variables such as the **year of death, cause of death, age group, gender**, and **number of deaths**.
-It classifies individuals into age categories spanning 20 years, starting from **"\<20"** years to **"60+"** years, with gender recorded as either **"Male"** or **"Female"**.
-The dataset enumerates the deaths annually from **2017** through **2023**, categorizing the causes into distinct classifications like **"Accident", "Drug Toxicity", "Suicide"**, among others.
+The dataset comprises variables such as the year of death, cause of death, age group, gender, and number of deaths.
+It classifies individuals into age categories spanning 20 years, starting from 20 years to 60+ years, with gender recorded as either Male or Female.
+The dataset enumerates the deaths annually from 2017 through 2023, categorizing the causes into distinct classifications like Accident, Drug Toxicity", Suicide, among others.
 
-Focusing on suicide-related fatalities, the dataset was refined to exclude entries marked as **"Unknown"** or **empty**.
+Focusing on suicide-related fatalities, the dataset was refined to exclude entries marked as Unknown or empty.
 This filtration led to the construction of two specialized datasets: one delineating the yearly suicide death toll segregated by gender and another by age group.
@@
 
-Table tbl-suicide-age-group presents the refined data, specifically spotlighting instances of suicide by age groups, and depicting the temporal and numerical specifics of each case.
+Table @tbl-suicide-age-group presents the refined data, specifically spotlighting instances of suicide by age groups, and depicting the temporal and numerical specifics of each case.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
 #| label: tbl-suicide-age-group
@@
 #| echo: false
+#| warning: false
+#| message: false
 
 # Loading the dataset
-suicide_by_age <- read_csv(here::here("data/analysis_data/age_group_deaths_data.csv"))
+suicide_by_age <- read_csv(here::here("data/analysis_data/age_group_deaths_data.csv"), show_col_types = FALSE)
 
-suicide_by_age %>%
-  select(Year, Age_group, Deaths) %>%
-  head(8) %>%
+# Aggregating data to show total counts for each age group
+suicide_summary <- suicide_by_age %>%
+  group_by(Age_group) %>%
+  summarise(Total_Counts = sum(Count)) %>%
+  arrange(desc(Total_Counts))
+
+# Creating a concise table
+suicide_summary %>%
   kable(
-    col.names = c("Year", "Age Group", "Deaths"),
+    col.names = c("Age_group", "Total Counts"),
     digits = 1,
@@
   ) %>%
-
-  # Ensuring the table maintains its designated position
   kable_styling(latex_options = "HOLD_position")
@@
 
-Table tbl-suicide-gender-group presents the refined data, specifically spotlighting instances of suicide by gender, and depicting the temporal and numerical specifics of each case.
+Table @tbl-suicide-gender-group presents the refined data, specifically spotlighting instances of suicide by gender, and depicting the temporal and numerical specifics of each case.
 
@@
 # Load the dataset containing gender-grouped suicide data
-gender_data <- read_csv(here::here("data/analysis_data/gender_deaths_data.csv"))
+gender_data <- read_csv(here::here("data/analysis_data/gender_group_deaths_data.csv"))
 
@@
 gender_data |> 
-  select(Year, Gender, Deaths) |>  # Select relevant columns
+  select(Year, sex, count) |>  # Select relevant columns
   slice(1:8) |>  
   kable(
-    col.names = c("Year", "Gender", "Deaths"),  
+    col.names = c("Year", "sex", "count"),  
     digits = 1,  
@@
 
-# Model
+# Model{#sec-model}
 
 ## Model set-up 
+This section elucidates the development of a logistic regression model, tailored to predict the likelihood of suicide among homeless individuals using demographic factors.
+Before delving into the model, we visually explore the relationship between suicide rates and demographic variables—age and gender—using ggplot2.
+
+We employ @graph-suicide-gender-group to create a comprehensive visual representation, showcasing the relationship between suicide rates and the key demographic variables, age and gender.
+This visualization aids in understanding the data distribution and any apparent trends that could influence the model.
+
+```{r}
+#| label: graph-suicide-gender-group
+#| fig-cap: Summary of Suicide Rates by Age and Gender
+#| echo: false
+#| warning: false
+#| message: false
+
+library(ggplot2)
+library(readr)
+
+cleaned_data <- read_csv(here::here("data/analysis_data/cleaned_data.csv"), show_col_types = FALSE)
+
+ggplot(cleaned_data, aes(x = Age_group, y = Count, fill = Gender)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  labs(x = "Age Group",
+       y = "Suicide Rate") +
+  theme_minimal() +
+  scale_fill_brewer(palette = "Set1")
+```
+
+The logistic regression model is designed to predict the probability of suicide cases among homeless individuals, utilizing sociodemographic factors such as age and gender.
+In a Bayesian logistic regression context, the model's setup can be represented with the following hierarchical structure:
+
+\begin{align}
+y_i | p_i &\sim \mbox{Bernoulli}(p_i) \\
+\log\left(\frac{p_i}{1 - p_i}\right) &= \alpha + \beta_1 \times \text{AgeGroup}_{i1} + \beta_2 \times \text{AgeGroup}_{i2} + \beta_3 \times \text{Gender}_{i} \\
+\alpha &\sim \mbox{Normal}(0, 2.5) \\
+\beta_j &\sim \mbox{Normal}(0, 2.5) \, \text{for } j = 1, 2, 3 \\
+\end{align}
+
+In this adapted model:
+
+- $y_i$ represents the binary outcome for each individual (i.e., the presence or absence of a suicide case).
+- $p_i$ is the probability of observing a suicide case for the ith individual, linked to the predictors through the logistic function.
+- The coefficients $\alpha$,$\beta_1$, $\beta_2$, and $\beta_3$ are assigned Normal prior distributions, reflecting our prior beliefs about these parameters' distributions before observing the data.
+The Normal priors are centered at 0 with a standard deviation of 2.5, indicating moderate certainty in the prior information.
+- The logit link function (log-odds) is the natural logarithm of the odds $p_i$ / $1 - p_i$ and linearly relates the predictors to the probability of the outcome.
+
+### Model justification 
+This Bayesian logistic regression framework not only estimates the parameters but also quantifies uncertainty in the estimates, providing a more comprehensive understanding of the model's predictions and the effects of the predictors.
+The use of Normal priors for the coefficients is a standard choice in Bayesian modeling, offering flexibility and conjugacy that facilitate both analytical and computational solutions.
+The hierarchical model structure allows for the incorporation of additional layers of complexity and variability, accommodating more nuanced relationships and potential hierarchical data structures.
 
@@
 
+# Discussion
+
 # References

### 2024-03-14 scripts/02-data_cleaning.R

@@ -9,7 +9,7 @@
 library(tidyverse)
 
 #### Clean data ####
-dataset <- read_csv("data/raw_data/Homeless_deaths_by_cause.csv")
+dataset <- read_csv(here::here("data/raw_data/Homeless_deaths_by_cause.csv"), show_col_types = FALSE)
 
 # Rename the year of death column
 names(dataset)[names(dataset) == 'Year of death'] <- 'Year_of_death'
@@ -26,50 +26,44 @@ head(cleaned_data, n=10)
 #### Save data ####
 write_csv(cleaned_data, "data/analysis_data/cleaned_data.csv")
 
-### Data for sex###
+### Data for gender###
 
-# Initialize a list to store the sum of deaths of Male
-Male_deaths_summary <- list()
-
-for (year in c(2017,2018,2019,2020,2021, 2022, 2023)) {
-  # Filter data for Male
-  year_data <- cleaned_data %>%
-    filter(Year_of_death == year, Gender == 'Male') %>%
-    select(Year_of_death, Gender, Count)
-  
-  print(head(year_data, n = 10))
-  
-  # Calculate the sum of deaths
-  Male_deaths_summary[[as.character(year)]] <- sum(year_data$Count)
-}
-
-# Initialize a list to store the sum of deaths of Female
-Female_deaths_summary <- list()
-
-for (year in c(2017,2018,2019,2020,2021, 2022, 2023)) {
-  # Filter data for Female
-  year_data <- cleaned_data %>%
-    filter(Year_of_death == year, Gender == 'Male') %>%
-    select(Year_of_death, Gender, Count)
-  
-  print(head(year_data, n = 10))
-  
-  # Calculate the sum of deaths
-  Female_deaths_summary[[as.character(year)]] <- sum(year_data$Count)
-}
-
-# Load the tibble package for creating tibbles
+library(dplyr)
 library(tibble)
+library(readr)
+
+# Define a function to summarize deaths by year and gender group
+summarize_deaths_by_gender <- function(Gender, data) {
+  deaths_summary <- numeric(length = length(2017:2023))
+  years <- 2017:2023
+  
+  for (i in seq_along(years)) {
+    year <- years[i]
+    year_data <- data %>%
+      filter(Year_of_death == year, Gender == !!Gender) %>%
+      select(Year_of_death, Gender, Count)
+    
+    print(head(year_data, n = 10))
+    
[…]

### 2024-03-15 paper/paper.qmd
Update

@@
 
-## Data source
+## Data Source
 
-The study was performed using data from the City of Toronto's **opendatatoronto** database portal, accessed through the 'opendatatoronto' package and processed using the statistical programming environment **R**.
-The **tidyverse** package facilitated the data and the **kableExtra** package was utilized for enhancing table presentations.
+The study was performed using data from the City of Toronto's database portal,@citeopendatatoronto, accessed through the 'opendatatoronto' package and processed using the statistical programming environment @citeR.
+The @citetidyverse package facilitated the data and the @rGgplot2, @rKnitr, @rReadr and @rTibble package was utilized for enhancing table presentations.
 Specific to this study, the LaTeX package was implemented in the R markdown setting to ensure stable positioning of figures and tables.
 
-### Data measurment
+## Data Measurment
 
@@
 #| label: tbl-suicide-age-group
-#| tbl-cap: Summary of Suicides by Age Group
+#| tbl-cap: Suicides Rates by Age Group
 #| echo: false
@@
 #| label: tbl-suicide-gender-group
-#| fig-cap: Summary of suicides Rates by Gender
+#| tbl-cap: Suicides Rates by Gender
 #| echo: false
@@
 #| message: false
-
-library(knitr)
 
@@
   select(Year, sex, count) |>  # Select relevant columns
-  slice(1:8) |>  
+  slice(1:14) |>  
   kable(
@@
 
-## Model set-up 
+## Model Set-up 
 This section elucidates the development of a logistic regression model, tailored to predict the likelihood of suicide among homeless individuals using demographic factors.
@@
 
-### Model justification 
+## Model Justification 
 This Bayesian logistic regression framework not only estimates the parameters but also quantifies uncertainty in the estimates, providing a more comprehensive understanding of the model's predictions and the effects of the predictors.
@@
 
-# Discussion
+The regression model we've employed is designed to explore the relationships between the count of reported suicide incidents and several predictor variables, specifically the year of death, age group, and gender, within Toronto's homeless population.
+This model is built using linear regression, a statistical method that attempts to model the relationship between a dependent variable (in this case, the count of suicides) and one or more independent variables (year, age group, and gender) by fitting a linear equation to the observed data.
+
+```{r}
+#| label: graph-model-one
+#| fig-cap: Residuals vs. Fitted Values
+#| echo: false
+#| warning: false
+#| message: false
+
+# Load necessary library
+library(ggplot2)
+
+# Assuming your dataset is loaded as 'cleaned_data' and your model is fitted as follows:
+model <- lm(Count ~ Year_of_death + Age_group + Gender, data = cleaned_data)
+
+residuals_plot <- ggplot(cleaned_data, aes(x = model$fitted.values, y = model$residuals)) +
+  geom_point() +
+  geom_hline(yintercept = 0, linetype = "dashed") +
+  labs(x = "Fitted Values", y = "Residuals")
+
+print(residuals_plot)
+```
+One of the primary diagnostic plots for assessing a linear regression model is the residuals vs. fitted values plot. To understand this, we first need to clarify what residuals and fitted values are:
+
+- **Fitted Values**: These are the estimated values of the dependent variable (suicide counts) produced by the regression model when it applies the estimated coefficients to the predictor variables.
+They represent the model's predictions.
+
+- **Residuals**: These are the differences between the observed values (actual suicide counts in the data) and the fitted values (the model's predictions).
+Essentially, a residual is an error in the prediction; it tells us how much the model has missed the actual value.
+
+```{r}
+#| label: graph-model-two
+#| fig-cap: Actual vs. Predicted Values
+#| echo: false
+#| warning: false
+#| message: false
+
+# Assuming your dataset is loaded as 'cleaned_data' and your model is fitted as follows:
+model <- lm(Count ~ Year_of_death + Age_group + Gender, data = cleaned_data)
+
+predictions <- predict(model, cleaned_data)
+actual_vs_predicted_plot <- ggplot(cleaned_data, aes(x = Count, y = predictions)) +
+  geom_point() +
+  geom_abline(intercept = 0, slope = 1, col = "red") +
+  labs(x = "Actual Values", y = "Predicted Values")
+
+print(actual_vs_predicted_plot)
+```
+
+This figure is a fundamental visualization in regression analysis, aimed at comparing the observed actual data points with the values predicted by the model. Here's how it works and what it signifies:
+
+- **Actual Values**: These are the real observed values of the dependent variable from your dataset—in this case, the count of suicide occurrences within the studied population.
+
+- **Predicted Values**: Predicted values are generated by the model when the estimated regression coefficients are applied to the predictor variables.
+They represent the model's estimations or predictions of the dependent variable based on the relationships it has identified.
+
+# Discussion {#sec-model}
+## Overview of the Study
+
+This paper embarked on an analytical journey to uncover the intricate relationship between sociodemographic factors and the prevalence of suicide among Toronto's homeless population.
+Employing a robust regression model, we scrutinized the impact of age, gender, and temporal factors on suicide occurrences, unveiling nuanced patterns and trends that characterize this grave public health issue.
+Our model's meticulous analysis, supported by rigorous statistical methods, provided a comprehensive perspective on how demographic variables correlate with suicide rates in the context of homelessness.
+
+One pivotal insight gleaned from our study is the profound influence of age on suicide risks among the homeless.
+The differentiated age brackets revealed significant disparities, highlighting a particularly vulnerable segment of the population.
+This finding underscores the critical need for age-specific intervention strategies and support services, which could be pivotal in mitigating suicide risks among Toronto's homeless individuals.
+
+Furthermore, the study illuminated notable temporal trends in suicide rates, offering a window into how these patterns have evolved over the years.
+This temporal analysis is crucial, as it reflects the dynamic nature of the issue, influenced by varying societal, economic, and policy-driven factors.
+Recognizing these trends is essential for policymakers and healthcare providers to adapt their strategies effectively, ensuring that they remain relevant and impactful in addressing the needs of the homeless population.
+
+## Reflection on the Study
+
+While our study provides significant insights, it is not without its limitations.
+The reliance on retrospective data and the inherent constraints of a regression analysis model might have influenced the comprehensiveness of our findings.
+Potential biases in data reporting and the challenge of accurately capturing the multifaceted experiences of homelessness could affect the generalizability of our results.
+Additionally, the cross-sectional nature of our data limits our ability to infer causality between the examined sociodemographic factors and suicide occurrences.
+
+The path forward necessitates a multi-faceted approach to further explore and address this complex issue.
+Future research should consider longitudinal studies to better understand the causal relationships and long-term trends in homelessness and suicide.
 

### 2024-03-16 paper/paper.qmd
update

--- 2024-03-14 paper/paper.qmd
@@
 ---
-title: "An Analytical Exploration of Homelessness and Suicide in Toronto"
+title: "Identifying Risk: Predictive Model of Suicide Among Toronto's Homeless"
 author: 
@@
 date: today
-abstract: "This paper presents a study focused on addressing the critical issue of suicide among Toronto's homeless population, a significant public health challenge.
-We developed a risk stratification model using sociodemographic data to identify homeless individuals at high risk for suicide, employing a generalized linear model that incorporates factors like sex and age.
-The findings, derived from data collected by Toronto Public Health since January 2017, offer insights into the patterns and predictors of suicide attempts within this vulnerable group.
-The importance of this research lies in its potential to guide interventions, enabling more effective prevention strategies for suicide among the homeless, particularly in settings where immediate access to mental health professionals is limited."
+abstract: "This study unveils a novel risk stratification model to assess suicide risks among Toronto's homeless population, using sociodemographic data within a generalized linear model.
+Our analysis, grounded in data from Toronto Public Health, reveals critical patterns and predictors of suicidal behaviour, providing pivotal insights for intervention strategies.
+The findings highlight the urgent need for targeted preventive measures, emphasizing the model's potential in shaping public health policies and social services, thereby contributing significantly to mitigating suicide risks in marginalized communities."
 format:
@@
 This paper delves into the development of a risk stratification model designed to identify homeless individuals at an elevated risk of suicide.
-By integrating sociodemographic data within a generalized linear model, focusing on sex and age as primary factors, our research aims to provide a predictive tool for healthcare providers and social services.
+By integrating sociodemographic data within a generalized linear model (Bayesian logistic regression model) focusing on sex and age as primary factors, our research aims to provide a predictive tool for healthcare providers and social services.
 The motivation behind this study stems from the critical gap in existing research regarding effective, data-driven strategies for suicide prevention in the homeless community.
@@
 The outcomes of our study revealed significant patterns and predictors that can enhance the effectiveness of suicide prevention strategies, offering a new perspective on addressing this public health challenge.
-The significance of our findings extends beyond academic interest; they offer practical insights for policymakers, healthcare providers, and social workers who are directly involved in supporting the homeless population.
 By identifying high-risk individuals, our model can inform targeted interventions, potentially saving lives and allocating resources more efficiently.
 
-The structure of the paper is organized as follows: The data section delves into the broader context of the dataset, emphasizing the crucial aspects of measurement relevant to our study.
-The results section is comprehensive, encompassing various illustrative tables and graphs alongside detailed statistical analyses.
-Subsequently, the model section elucidates the setup and justification of our chosen modelling approach, ensuring clarity and transparency of our methodology.
-Lastly, the discussion articulates the study's contributions to our understanding of homelessness and suicide and outlines directions for future research, ensuring a thorough contemplation of the study's broader implications.
+The structure of the paper is organized as follows: @sec-data (Data) delves into the broader context of the dataset, emphasizing the crucial aspects of measurement relevant to our study.
+@sec-model (Model) elucidates the setup and justification of our chosen logistic regression model, ensuring clarity and transparency of our methodology.
+Subsequently, @sec-results (Results) presents the coefficients of the regression model alongside detailed statistical analyses.
+Lastly, @sec-discussion (Discussion) articulates the study's contributions to our understanding of homelessness and suicide and outlines directions for future research, ensuring a thorough contemplation of the study's broader implications.
 
@@
 
-## Data source
+## Data Source
 
-The study was performed using data from the City of Toronto's **opendatatoronto** database portal, accessed through the 'opendatatoronto' package and processed using the statistical programming environment **R**.
-The **tidyverse** package facilitated the data and the **kableExtra** package was utilized for enhancing table presentations.
-Specific to this study, the LaTeX package was implemented in the R markdown setting to ensure stable positioning of figures and tables.
+The study was performed using data from the City of Toronto's database portal,[@citeopendatatoronto], accessed through the 'opendatatoronto' package and processed using the statistical programming environment R [@citeR].
+The tidyverse [@citetidyverse] package facilitated the data and the ggplot [@rGgplot2], knitr [@rKnitr], readr [@rReadr] and tibble [@rTibble] package was utilized for enhancing table presentations.
+Specific to this study, the kableExtra [@citekableExtra] package was implemented in the R markdown setting to ensure stable positioning of figures and tables.
 
-### Data measurment
+## Data Measurment
 
@@
 
-Table @tbl-suicide-age-group presents the refined data, specifically spotlighting instances of suicide by age groups, and depicting the temporal and numerical specifics of each case.
+## Data Summary
+
+Table @tbl-suicide-age-group presents the refined data, specifically spotlighting instances of suicide by age groups, and depicting the numerical specifics of each case.
 
@@
 
-We employ @graph-suicide-gender-group to create a comprehensive visual representation, showcasing the relationship between suicide rates and the key demographic variables, age and gender.
+We employ @fig-suicide-gender-group to create a comprehensive visual representation, showcasing the relationship between suicide rates and the key demographic variables, age and gender.
 This visualization aids in understanding the data distribution and any apparent trends that could influence the model.
@@
 ```{r}
-#| label: graph-suicide-gender-group
+#| label: fig-suicide-gender-group
 #| fig-cap: Summary of Suicide Rates by Age and Gender
@@
 
+## Model Equation 
 The logistic regression model is designed to predict the probability of suicide cases among homeless individuals, utilizing sociodemographic factors such as age and gender.
@@
 - $y_i$ represents the binary outcome for each individual (i.e., the presence or absence of a suicide case).
-- $p_i$ is the probability of observing a suicide case for the ith individual, linked to the predictors through the logistic function.
+- $p_i$ is the probability of observing a suicide case for the individual, linked to the predictors through the logistic function.
 - The coefficients $\alpha$,$\beta_1$, $\beta_2$, and $\beta_3$ are assigned Normal prior distributions, reflecting our prior beliefs about these parameters' distributions before observing the data.
@@
 
-### Model justification 
-This Bayesian logistic regression framework not only estimates the parameters but also quantifies uncertainty in the estimates, providing a more comprehensive understanding of the model's predictions and the effects of the predictors.
-The use of Normal priors for the coefficients is a standard choice in Bayesian modeling, offering flexibility and conjugacy that facilitate both analytical and computational solutions.
-The hierarchical model structure allows for the incorporation of additional layers of complexity and variability, accommodating more nuanced relationships and potential hierarchical data structures.
+## Model Justification 
+The Bayesian logistic regression model is adeptly suited for the paper's objective of creating a risk stratification tool to identify homeless individuals at elevated suicide risk.
+It judiciously integrates sociodemographic factors like age and gender, utilizing a Bayesian framework to incorporate prior knowledge and present data, thus offering a robust predictive analysis.
+The model's hierarchical structure and the inclusion of Normal priors for coefficients ensure a nuanced analysis, balancing prior beliefs with empirical data, crucial for strategizing suicide prevention efforts among the homeless.
 
-# Results
+# Results {#sec-results}
 
-# Discussion
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+# Retrieve the Bayesian linear model from the saved .rds file
+bayesian_model <- readRDS(file = here("models", "bayesian_risk_stratification_model.rds"))
+```
+
+## Model Coefficeints Interpretation 
+@tbl-residuals-table provides a detailed summary of the estimated effects of age groups and gender on the suicide risk among homeless individuals, as derived from the Bayesian regression analysis.
+Each row in the table represents a different predictor in the model, with the coefficients indicating the magnitude and direction of the association between that predictor and the observed counts of suicide cases.
+For instance, if the coefficient for a certain age group is positive and statistically significant, it suggests that individuals in this age group have a higher risk of suicide compared to the baseline group, after controlling for other factors in the model.
+Conversely, a negative coefficient would suggest a lower risk.
+Similarly, the gender coefficients shed light on the differential risk of suicide between genders.
+
+```{r}
+#| label: tbl-residuals-table
+#| tbl-cap: Summary of Residuals for the Bayesian Model
+#| echo: false
+#| warning: false
+#| message: false
+
+library(knitr)
+library(kableExtra)
+
+# Extract residuals from the Bayesian model
+residuals_data <- residuals(bayesian_model)
+
+# Select the first 10 observations
+residuals_subset <- head(residuals_data, 10)
+
+# Create a data frame from the subset of residuals
+residuals_df <- data.frame(
+  Observation = seq_along(residuals_subset),
+  Residual = residuals_subset
+)
+
+residuals_df$Residual <- round(residuals_df$Residual, 3)
+
+residuals_table <- kable(residuals_df,
+                         col.names = c("Observation", "Residual"), align = 'c') %>%
+  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE,
+                position = "center")
+
+residuals_table
+```
+
+@tbl-residuals-table is crucial for justifying the thesis of the paper as it provides empirical evidence supporting the hypothesis that certain age groups or genders are at a higher risk of suicide among homeless individuals.
+By quantifying these associations, the model helps in identifying high-risk subgroups within the homeless population, thereby offering insights that can inform targeted interventions and preventive strategies.
+The Bayesian framework further enriches this analysis by incorporating prior knowledge and uncertainty into the estimation process, thereby enhancing the robustness and interpretability of the findings.
+
+## Model Visualization 
+@fig-suicide-Cases effectively illustrates the relationship between demographic factors—specifically age groups and gender—and the probability of suicide among homeless individuals.
+This graphical representation is pivotal in understanding how these variables interplay to influence suicide risk.
+
+Each point on the plot corresponds to a specific age group and is color-coded to represent gender, showcasing a nuanced view of how suicide counts vary across different demographic segments.
+The size of each point is proportional to the count of suicide cases, providing an immediate visual cue to the relative magnitude of suicide risk within each category.
+This size differentiation helps in identifying which combinations of age group and gender have higher or lower counts of suicide incidents, thereby indicating potential risk stratification within the homeless population.
+
+```{r}
+#| label: fig-suicide-Cases
+#| fig-cap: Suicide Cases by Age Group and Gender
+#| echo: false
+#| warning: false
+#| message: false
+
+library(ggplot2)
+library(forcats)  # For reordering factors
+
+cleaned_data$Age_group <- fct_reorder(cleaned_data$Age_group, cleaned_data$Count)
+
+# Generating scatter plot
+advanced_plot <- ggplot(cleaned_data, aes(x = Age_group, y = Count, color = Gender)) +
+  geom_point(aes(size = Count), alpha = 0.6) +  # Points sized by Count
+  geom_smooth(method = "lm", se = FALSE, color = "gray") +  # Add a linear model fit line
+  scale_size_continuous(name = "Count of Cases", range = c(3, 10)) +  # Adjust the size scale
+  labs(
+    x = "Age Group",
+    y = "Count of Suicide Cases"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(size = 16),
+    axis.title.x = element_text(size = 14),
+    axis.title.y = element_text(size = 14),
+    legend.title = element_text(size = 12),
+    legend.text = element_text(size = 10)
+  ) +
+  scale_color_manual(values = c("blue", "pink"))  # Assign manual colors for gender
+
+print(advanced_plot)
+```
+
+# Discussion {#sec-discussion}
+## Overview of the Study
+
+This research delved into the complex interplay between sociodemographic factors and the incidence of suicide within the homeless population of Toronto.
+Through a detailed regression analysis, we explored how variables such as age, gender, and time-related elements influence suicide rates, uncovering specific patterns and insights pivotal to understanding this critical public health concern.
+The precision of our model, grounded in solid statistical practices, offered an in-depth view of the demographic determinants of suicide risk among the homeless.
+
+A key discovery of our analysis was the pronounced impact of gender and age on suicide risk, with the model indicating that males in the 40-59 age group are at a higher risk compared to other demographics.
+This group emerged as particularly susceptible, signaling an urgent need for targeted interventions and support mechanisms.
+Additionally, the model revealed that younger males, specifically those aged 20-39, also face a heightened risk of suicide attempts, pointing towards the necessity for early preventive measures tailored to this subgroup.
+
+Contrastingly, the results showed that females generally have lower instances of suicide attempts, a finding that, while reassuring, also calls for a nuanced understanding of gender-specific vulnerabilities and protective factors in the context of homelessness.
+
+Moreover, our analysis shed light on the changing trends in suicide rates over time, providing insights into the fluctuating dynamics of this issue.
+Such temporal insights are vital for adapting intervention strategies and policy formulations to the evolving landscape of societal, economic, and health-related influences affecting the homeless population.
+
+These findings not only emphasize the importance of considering age and gender in suicide prevention strategies but also highlight the necessity for dynamic, evidence-based approaches to support Toronto's homeless community, ultimately aiming to reduce the prevalence of this tragic outcome.
+
+## Reflection on the Study
+
+Our investigation yields valuable findings, yet it is important to acknowledge its limitations.
+The study's dependency on historical data, coupled with the intrinsic limitations associated with regression analysis, may have impacted the depth of our insights.
+The possibility of data reporting biases, along with the difficulty in fully encapsulating the complex realities of homelessness, might limit how broadly our conclusions can be applied.
+Moreover, the snapshot nature of our data constrains our capacity to establish causative links between the sociodemographic variables studied and the instances of suicide.
+
+Looking ahead, addressing this intricate issue demands a comprehensive strategy.
+It's imperative for upcoming research to incorporate longitudinal designs, enabling a clearer understanding of the causative dynamics and enduring patterns connecting homelessness with suicide risks.
+This approach will enhance our grasp of the temporal aspects of these factors and their long-term consequences, paving the way for more effective interventions and policy-making.
 

### 2024-03-16 scripts/04-model.R

@@ -7,31 +7,31 @@
 
 
 #### Workspace setup ####
-library(tidyverse)
+library(readr)
 library(rstanarm)
+library(here)
 
 #### Read data ####
-modeling_data <- read_csv("data/analysis_data/cleaned_data.csv")
+cleaned_data <- read_csv(here("data", "analysis_data", "cleaned_data.csv"))
 
-### Model data ####
-first_model <- stan_glm(
-  formula = Count ~ Gender + Age_group,
-  data = modeling_data,
-  family = gaussian(),
-  prior = normal(0, 23, autoscale = TRUE),
-  prior_intercept = normal(0, 1, autoscale = TRUE),
-  prior_aux = exponential(0.33, autoscale = TRUE),
-  seed = 80
+# Ensure reproducibility
+set.seed(2024)
+
+#### Model data ####
+bayesian_model <- stan_glm(
+  Count ~ Age_group + Gender, 
+  data = cleaned_data,
+  family = poisson(link = "log"),
+  seed = 2024 # Set the seed for reproducibility
 )
 
+# Summary of the model to check for immediate red flags
+summary(bayesian_model)
 
 #### Save model ####
-# Save the constructed model for future reference
+# Save the Bayesian model as an .rds file to the 'models' folder
 saveRDS(
-  first_model,
-  file = "models/first_model.rds"
+  bayesian_model,
+  file = here("models", "bayesian_risk_stratification_model.rds")
 )
-
-# Notification of script completion
-message("first_model has been successfully created and saved.")
 
